---
title: "Analysis_pipeline1"
author: "Jana Fischer"
date: "1/30/2020"
output: html_document
---

Clean single-cell data and perform first analysis steps, including distinguishing tumor and stromal cells and dimensionality reduction.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(data.table)
library(SingleCellExperiment)
library(ggplot2)
library(mclust)
library(RColorBrewer)
library(scater)
library(plyr)
library(viridis)
library(ggrastr)
library(ComplexHeatmap)
library(circlize)

```

Read SCE with single-cell data and clean single-cell data.
```{r,eval = FALSE}
#Set global seed
set.seed(3)

#Read in single cell experiment created in CP_to_SCE.Rmd
sce = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE.rds')

#Exclude second Iridium channel (don't need it for single-cell analysis)
sce = sce[rownames(sce)[!rownames(sce) %in% c("Ir191_Iridium")],]

#Flag very small or very big cells as potential missegmentations (excluded from most analyses)
sce$sizeflag[sce$AreaShape_Area >= 700] = 1
sce$sizeflag[sce$AreaShape_Area <= 10] = 1
sce$sizeflag[is.na(sce$sizeflag)] = 0

#Exclude split image with two half cores (unclear from which TMA location)
sce = subset(sce,,core != "ZTMA21.3_s0_p3_r10_a10")

#Exclude much too small images (created due to machine interruprions)
sce = subset(sce,,core != "ZTMA26.1_s0_p3_r38_a38") #narrow stripe
sce = subset(sce,,core != "ZTMA21.1_s0_p3_r162_a162") #tiny tile

#Save batch nr as metadata (not just which TMA but also which acquisition round)
sce$ArrayNr_batch = sub("_s.*", "",sce$core)

#Check for bad stains (looking at images and their cell types as well as metadata showed no systematic badly stained cores)
d = as.data.table(t(counts(sce)))
d$id = colnames(sce)
d$core = sce$core
d$meta = sce$Metadata_Description #holds info about TMA location and split cores
d[,meanIr := mean(Ir193_Iridium),by = "core"]
d[,meanHist := mean(`In113_Histone H3`),by = "core"]
d$comb_nuc = d$meanHist + d$meanIr
d = unique(d[,c("comb_nuc","core","meta")])

ggplot(d,aes(x = comb_nuc))+geom_density()

#Images visually inspected
quant = quantile(d$comb_nuc,0.05)
d[comb_nuc < quant]

```

Run a granular PG on all cells and a gaussian mixture model on panCK to distinguish tumor vs stromal cells.
```{r,eval = FALSE}
#Set global seed
set.seed(3)

#Run PG
rpheno_out = cytofkit::Rphenograph(t(counts(sce)), k = 50, seed = 3, approx=T)
sce$PG_clusters = as.factor(igraph::membership(rpheno_out))

#Gaussian mixture model to distinguish tumor/stroma based on panCK expression
mcl.model <- densityMclust(t(counts(sce))[,"Lu175_Keratin Epithelial"], 2)
sce$GMM_class = mcl.model$classification

#Save another SCE with PG output included (so this slow step doesn't have to be rerun)
saveRDS(sce, file = '/mnt/ZTMA_21_25_26/analysis/SCE_PG.rds', ascii = FALSE)

```

Read saved SCE with above clustering info already contained and define frequently used colormaps
```{r}
#In case this wasn't already set
set.seed(3)

#Read in sce with PG info already contained
sce = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE_PG.rds')

#Define general color map of distinguishable colors
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_vector = unique(col_vector)

#Define colormaps for the frequently used metadata
#Molecular subtype
cols_mol = unique(as.data.table(sce$Mol_Signature))
cols_mol$col = c("pink","violet","black","turquoise","orange","purple","white")
names(cols_mol)[1] = "Mol_Signature"

#Pathology grade
cols_grade = unique(as.data.table(sce$Grade))
cols_grade$col = c("red","green","blue","white","black")
names(cols_grade)[1] = "Grade"

#Nodal status
cols_pN_4gr = unique(as.data.table(sce$pN_4gr))
cols_pN_4gr$col = brewer.pal(nrow(cols_pN_4gr), "OrRd")
names(cols_pN_4gr)[1] = "pN_4gr"
cols_pN_4gr$col[is.na(cols_pN_4gr$pN_4gr)] = "black"
cols_pN_4gr$col[cols_pN_4gr$pN_4gr == "7"] = "grey"

#Tissue type
cols_tissue = unique(as.data.table(sce$TissueType))
names(cols_tissue) = "TissueType"
cols_tissue$col = c("steelblue1","orange","black","purple","darkgreen","yellow","red")

#Write out color maps for following scripts
fwrite(cols_mol,"/mnt/ZTMA_21_25_26/analysis/cols_mol.csv",col.names = T)
fwrite(cols_grade,"/mnt/ZTMA_21_25_26/analysis/cols_grade.csv",col.names = T)
fwrite(cols_pN_4gr,"/mnt/ZTMA_21_25_26/analysis/cols_pN_4gr.csv",col.names = T)
fwrite(cols_tissue,"/mnt/ZTMA_21_25_26/analysis/cols_tissue",col.names = T)

```

Plot gaussian mixture model
```{r}
d = data.table(gMM = sce$GMM_class,panCK = t(assay(sce["Lu175_Keratin Epithelial",],type = "raw")))
d$`panCK.Lu175_Keratin Epithelial` = asinh(d$`panCK.Lu175_Keratin Epithelial`)

ggplot(d,aes(x=`panCK.Lu175_Keratin Epithelial`,fill = as.factor(gMM)))+geom_density()

```

Run t-SNE and UMAP dimensionality reductions
```{r,eval = F}
#In case this wasn't already set
set.seed(3)

#Register how many cores to use for parallelization
require(doParallel)
options('mc.cores' = 20)
registerDoParallel(20)

#Subsample 10% of cells from each core to save time and not get super crowded dimensionality reduction plots
sub = colData(sce)[c('core')]
sub$id = rownames(sub)
sub = as.data.table(sub)
sub = ddply(sub,.(core),function(x) x[sample(nrow(x),nrow(x)/10, replace = F),])
sce_sub = sce[,sub$id]

#Run umap and tsne and save output directly into subsampled SCE
sce_sub = runUMAP(sce_sub, exprs_values = "counts", external_neighbors=TRUE, BPPARAM = MulticoreParam())
sce_sub = runTSNE(sce_sub, exprs_values = "counts", external_neighbors=TRUE, BPPARAM = MulticoreParam())

#Save subsampled SCE including the dimensionality reduction info so this step doesn't have to be repeated
saveRDS(sce_sub, file = '/mnt/ZTMA_21_25_26/analysis/SCE_sub_dimRed.rds', ascii = FALSE)
```

Plot t-SNEs and UMAPs
```{r}
#Read in subsampeled sce with dim red info already contained
sce_sub = readRDS(file = '/mnt/ZTMA_21_25_26/analysis/SCE_sub_dimRed.rds')

#Add dummy variable consisting of all same entries because the default shape of plotReducedDim is ugly
sce_sub$dummy = 1

#Order factors according to color order
sce_sub$Mol_Signature = factor(sce_sub$Mol_Signature, levels = cols_mol$Mol_Signature)
sce_sub$Grade = factor(sce_sub$Grade, levels = cols_grade$Grade)
sce_sub$pN_4gr = factor(sce_sub$pN_4gr, levels = cols_pN_4gr$pN_4gr)
sce_sub$TissueType = factor(sce_sub$TissueType, levels = cols_tissue$TissueType)

#UMAPs
red = as.data.table(reducedDims(sce_sub)$UMAP)
shuffle = sample(nrow(red),replace = F) #shuffle points so all colors visible

#Plot density of points on UMAP
ggplot(red, aes(x=V1, y=V2))+
  stat_bin2d(aes(color=..count..), bins=300, geom='point', size=1.4, fill=1)+
  scale_color_viridis(trans = "log10", option='inferno')

#Tissue type
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$TissueType[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = cols_tissue[!is.na(TissueType)]$col)+
  theme_minimal()

#PID
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$PID[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = rep(col_vector,ceiling(length(unique(sce_sub$PID))/length(col_vector))))+
  theme_minimal()+ theme(legend.position = "none")

#Batch
TMAcols = c(colors()[grep("steelblue",colors())][1:4],
            colors()[grep("purple",colors())][1:4],
            colors()[grep("steelblue",colors())][5:7])

ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$ArrayNr_batch[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = TMAcols)+
  theme_minimal()

#Slide
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$ArrayNr[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = col_vector)+
  theme_minimal()

#Grade
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$Grade[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = cols_grade[!is.na(Grade)]$col)+
  theme_minimal()

#Molecular subtype
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$Mol_Signature[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = cols_mol$col)+
  theme_minimal()

#GMM class
ggplot(red[shuffle,],aes(x=V1, y=V2,color = as.factor(sce_sub$class[shuffle])))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = col_vector)

#Size flag
ggplot(red[shuffle,],aes(x=V1, y=V2,color = as.factor(sce_sub$sizeflag[shuffle])))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = col_vector)+
  theme_minimal()

#Plot UMAPs overlayed with expression values
par(mfrow=c(ceiling(length(rownames(sce_sub))/3),3))
lapply(rownames(sce_sub),function(x){plotReducedDim(sce_sub, dimred = "UMAP", shape_by = "dummy", by_exprs_values = "counts", colour_by = x)})


#t-SNEs
red = as.data.table(reducedDims(sce_sub)$TSNE)
shuffle = sample(nrow(red),replace = F) #shuffle points so all colors visible

#Plot density of points on tSNE
ggplot(red, aes(x=V1, y=V2))+
  stat_bin2d(aes(color=..count..), bins=300, geom='point', size=1.4, fill=1)+
  scale_color_viridis(trans = "log10", option='inferno')

#Tissuetype
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$TissueType[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = cols_tissue[!is.na(TissueType)]$col)+
  theme_minimal()

#PID
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$PID[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = rep(col_vector,ceiling(length(unique(sce_sub$PID))/length(col_vector))))+
  theme_minimal()+ theme(legend.position = "none")

#Batch
TMAcols = c(colors()[grep("steelblue",colors())][1:4],
            colors()[grep("purple",colors())][1:4],
            colors()[grep("steelblue",colors())][5:7])

ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$ArrayNr_batch[shuffle]))+
  geom_point_rast(size = 0.2,alpha = 0.5)+
  scale_color_manual(values = TMAcols)+
  theme_minimal()

#Slide
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$ArrayNr[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = col_vector)+
  theme_minimal()

#Grade
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$Grade[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = cols_grade[!is.na(Grade)]$col)+
  theme_minimal()

#Clinical type
ggplot(red[shuffle,],aes(x=V1, y=V2,color = sce_sub$Mol_Signature[shuffle]))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = cols_mol$col)+
  theme_minimal()

#GMM class
ggplot(red[shuffle,],aes(x=V1, y=V2,color = as.factor(sce_sub$class[shuffle])))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = col_vector)

#Size flag
ggplot(red[shuffle,],aes(x=V1, y=V2,color = as.factor(sce_sub$sizeflag[shuffle])))+
  geom_point_rast(size = 0.2)+
  scale_color_manual(values = col_vector)+
  theme_minimal()

#Plot t-SNEs overlayed with marker expression values
par(mfrow=c(ceiling(length(rownames(sce_sub))/3),3))
lapply(rownames(sce_sub),function(x){plotTSNE(sce_sub, shape_by = "dummy", by_exprs_values = "counts", colour_by = x)})

```

PCA
```{r}
#In case this wasn't already set
set.seed(3)

#Run PCA
sce_sub = runPCA(sce_sub, exprs_values = "counts")
dim(reducedDim(sce_sub, "PCA"))
plotReducedDim(subset(sce_sub,,TissueType %in% c("primary tumor","lymph node mts")), dimred = "PCA", colour_by = "TissueType")

#t-SNEs
red = as.data.table(reducedDims(subset(sce_sub,,TissueType %in% c("primary tumor","lymph node mts")))$PCA)
tistype = subset(sce_sub,,TissueType %in% c("primary tumor","lymph node mts"))$TissueType
shuffle = sample(ncol(subset(sce_sub,,TissueType %in% c("primary tumor","lymph node mts"))),replace = F) #shuffle points so all colors visible

ggplot(red[shuffle,],aes(x=PC1, y=PC2,color = tistype[shuffle]))+
  geom_point_rast(size = 1,alpha = 0.5)+
  scale_color_manual(values = c("steelblue1","purple"))+
  theme_minimal()

```

Heatmap of PG clusters
```{r}
#Marker heatmap
sum = sumCountsAcrossCells(sce,average = T,sce$PG_clusters, exprs_values = "counts")
p_dat = scale(t(sum))
p_dat[p_dat > 3] =3
p_dat[p_dat < -3] =-3

#Spatial feature heatmap
spatial = as.data.table(colData(sce)[c('Neighbors_NumberOfNeighbors_4','Neighbors_PercentTouching_4', "AreaShape_Area", "PG_clusters")])
spatial = aggregate(spatial[,-"PG_clusters"],by = list(PG_clusters = spatial$PG_clusters),FUN = mean)
rnames = spatial$PG_clusters
spatial = as.matrix(spatial[,colnames(spatial)[colnames(spatial) != "PG_clusters"]])
rownames(spatial) = rnames
spatial = scale(spatial)

#Indicate number of tumor/stromal cells per PG cluster as identified by gaussian mixture model on panCK
nr_tum = table(colData(sce)[c('PG_clusters','GMM_class')])
fr_tum = as.data.table(nr_tum)
fr_tum[,fr := N/sum(N),by = "PG_clusters"]
fr_tum = dcast.data.table(fr_tum,"PG_clusters ~ GMM_class", value.var = "fr")
rnames = fr_tum$PG_clusters
fr_tum=as.matrix(fr_tum[,-"PG_clusters"])
rownames(fr_tum) = rnames
fr_tum = fr_tum[order(as.numeric(rownames(fr_tum))),]

#Plot heatmap
h = Heatmap(p_dat, name = "markers",column_title = "markers", km = 1, col = colorRamp2(c(-2,0, 2), c('blue',"white", "red")), show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2")+
  Heatmap(spatial, name = "spatial",column_title = "spatial", km = 1, col = colorRamp2(c(-2,0, 2), c('blue',"white", "red")), show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2")+
  rowAnnotation(axis_reverse = anno_barplot(fr_tum, width = unit(10, "mm")),gp = gpar(fill = 1:2, col = 1:2))+
  rowAnnotation(axis_reverse = anno_barplot(nr_tum, width = unit(20, "mm")),gp = gpar(fill = 1:2, col = 1:2))+
  rowAnnotation(rn = anno_text(rownames(p_dat)))
h

#Tumor clusters
tumor = as.numeric(as.character(unique(sce$PG_clusters)))[!as.numeric(as.character(unique(sce$PG_clusters))) %in% c(60,59,126,115,82,26,16,125,11,70,36,103,44,42,61,34)]
tumor_bar = data.table(cluster = rownames(p_dat))
tumor_bar$epi = 0
tumor_bar[cluster %in% tumor,epi := 1]

#50% stromal
stroma = data.table(frac_stroma = fr_tum[,1])
stroma$epi = 1
stroma[frac_stroma >0.5, epi := 0]

#Mark tumor clusters in heatmap
h = Heatmap(p_dat, name = "markers",column_title = "markers", km = 1, col = colorRamp2(c(-2,0, 2), c('blue',"white", "red")), show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2")+
  Heatmap(spatial, name = "spatial",column_title = "spatial", km = 1, col = colorRamp2(c(-2,0, 2), c('blue',"white", "red")), show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2")+
  rowAnnotation(axis_reverse = anno_barplot(fr_tum, width = unit(10, "mm")),gp = gpar(fill = 1:2, col = 1:2))+
  rowAnnotation(axis_reverse = anno_barplot(nr_tum, width = unit(20, "mm")),gp = gpar(fill = 1:2, col = 1:2))+
  Heatmap(tumor_bar$epi, name = "cols", show_row_names = F, width = unit(10, "mm"), col = c("green","red"))+
  Heatmap(stroma$epi, name = "cols", show_row_names = F, width = unit(10, "mm"), col = c("green","red"))+
  rowAnnotation(rn = anno_text(rownames(p_dat)))
h
```


