---
title: "Analysis_pipeline5_revision_prep"
author: "Jana Fischer"
date: "26 1 2021"
output: html_document
---

Preparations for potential reviewer questions. Used R4.0 for this.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include = F}
library(Rphenograph)
library(igraph)
library(scater)
library(ComplexHeatmap)
library(circlize)
library(ggalluvial)
library(patchwork)
library(edgeR)
library(diffcyt)
library(data.table)
```

Cluster cells only based on ER, PR, HER2, Ki-67 (clinical markers) to see how much variation between primary and LN met there is then
```{r, eval = F}
#Read tumor single-cell data
sce_t = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE_T.rds')
clin_channels = c("Gd156_Estrogen Receptor Alpha","Gd158_Progesterone Receptor A/B","Er168_Ki-67","Eu151_c-erbB-2 - Her2")

#Cluster
set.seed(3)
rpheno_out = Rphenograph(t(counts(sce_t[clin_channels,])), k = 50)
sce_t$clin_clustering = as.factor(igraph::membership(rpheno_out[[2]]))

# #Save out
# #Save tumor SCE with clinical PG info
# saveRDS(sce_t, file = '/home/rstudio/mnt/ZTMA_21_25_26/analysis/SCE_T_clin_clustering.rds', ascii = FALSE)
```

Read in clustered data (saved out above) and plot heatmap
```{r}
#Read
sce_t = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE_T_clin_clustering.rds')

#Heatmap of markers used for clustering (don't use KI67 for metaclustering)
sum = sumCountsAcrossCells(sce_t[c("Gd156_Estrogen Receptor Alpha","Gd158_Progesterone Receptor A/B","Eu151_c-erbB-2 - Her2"),],average = T,colData(sce_t[c("Gd156_Estrogen Receptor Alpha","Gd158_Progesterone Receptor A/B","Eu151_c-erbB-2 - Her2"),])$clin_clustering, exprs_values = "counts")
#p_dat = scale(t(assay(sum))) #R4.0
p_dat = scale(t(sum))
p_dat[p_dat > 3] =3
p_dat[p_dat < -3] =-3

#Plot heatmap and split into 12 metaclusters
h_clusters_clin = Heatmap(p_dat, name = "clustered",column_title = "clustering markers", km = 1, col = colorRamp2(c(-2,0, 2), c('blue',"white", "red")),
            show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2",split = 12)
h_clusters_clin

#Extract metaclusters based on clustering of heatmap
metaclusters = lapply(row_order(h_clusters_clin), function(x){rownames(p_dat)[x]})
names(metaclusters) = 1:12
n = sapply(1:12,function(x){rep(x,length(metaclusters[[x]]))})
metaclusters = data.table(unlist(metaclusters))
metaclusters$meta = unlist(n)
names(metaclusters)[1] = 'clin_clustering'

#Manually metacluster some more into 9 clusters and name clusters
metaclusters[meta %in% c(2:3),meta := 2]
metaclusters[meta %in% c(4),meta := 3]
metaclusters[meta %in% c(5),meta := 4]
metaclusters[meta %in% c(6),meta := 5]
metaclusters[meta %in% c(7:8),meta := 6]
metaclusters[meta %in% c(9),meta := 7]
metaclusters[meta %in% c(10),meta := 8]
metaclusters[meta %in% c(11:12),meta := 9]
metaclusters[meta == 1, meta_name := "TNBC"]
metaclusters[meta == 2, meta_name := "PR+ER-"]
metaclusters[meta == 3, meta_name := "ERlowPR-"]
metaclusters[meta == 4, meta_name := "Luminallow/TNBC"]
metaclusters[meta == 5, meta_name := "HER2+HR-"]
metaclusters[meta == 6, meta_name := "HER2hiHR-"]
metaclusters[meta == 7, meta_name := "PRhiHER2lo"]
metaclusters[meta == 8, meta_name := "HR+HER2lo/+"]
metaclusters[meta == 9, meta_name := "ERhiHER2lo"]

#Add new clustering to sce_t
add = as.data.table(colData(sce_t)[,c("clin_clustering")])
add$id = colnames(sce_t)
names(add) = c("clin_clustering","id")
add = merge(add,metaclusters[,-"meta"],by = "clin_clustering",all.x = T)
add = add[order(match(id,colnames(sce_t))),]
sce_t$clin_meta_name = add$meta_name

#Heatmap of meta clusters
sum = sumCountsAcrossCells(sce_t[c("Gd156_Estrogen Receptor Alpha","Gd158_Progesterone Receptor A/B","Eu151_c-erbB-2 - Her2"),],average = T,colData(sce_t[c("Gd156_Estrogen Receptor Alpha","Gd158_Progesterone Receptor A/B","Eu151_c-erbB-2 - Her2"),])$clin_meta_name, exprs_values = "counts")
#p_dat = scale(t(assay(sum))) #R4.0
p_dat = scale(t(sum))
p_dat[p_dat > 3] =3
p_dat[p_dat < -3] =-3

#Plot
h_clusters_clin_meta = Heatmap(p_dat, name = "clustered",column_title = "clustering markers", km = 1, col = colorRamp2(c(-2,0, 2), c('blue',"white", "red")),
            show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2")
h_clusters_clin_meta
meta_order = rownames(p_dat)[row_order(h_clusters_clin_meta)]
```

FlowDiagram primary LN met
```{r}
#On most abundant cluster in LN met
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
#Exclude shitty images and cells
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
met[,test := .N,by = "PID"]
#met = met[test >=100,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('clin_meta_name','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

#Most abundant
m = apply(met[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(met[,-"PID"])[x]})

#Exclude ones with shared majority
PID_excl = met$PID[unlist(lapply(n,length)) > 1]
n = n[!unlist(lapply(n,length)) > 1]
maj_met = as.data.table(unlist(n))
names(maj_met) = "cluster_met"
maj_met$PID = met$PID[!met$PID %in% PID_excl]

#Most abundant cluster in primary tumor
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
#Exclude shitty images and cells
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
primary[,test := .N,by = "PID"]
#primary = primary[test >=100,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('clin_meta_name','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

m = apply(primary[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(primary[,-"PID"])[x]})

#Exclude ones with shared majority
PID_excl = primary$PID[unlist(lapply(n,length)) > 1]
n = n[!unlist(lapply(n,length)) > 1]
maj_prim = as.data.table(unlist(n))
names(maj_prim) = "cluster_prim"
maj_prim$PID = primary$PID[!primary$PID %in% PID_excl]

#Combine primary and LN met
comb = merge(maj_prim,maj_met,by = "PID")
comb[,Freq := .N ,by = c("cluster_prim","cluster_met")]
comb = unique(comb[,c("cluster_prim","cluster_met","Freq")])
comb$cluster_prim = factor(comb$cluster_prim,levels = meta_order)
comb$cluster_met = factor(comb$cluster_met,levels = meta_order)
clin_meta_cols = data.table(cluster = meta_order)
clin_meta_cols$cols = colors()[c(499,502,551,55,115,77,37,419,8)]

#Prepare color maps for ggalluvial
prim_cols = clin_meta_cols[cluster %in% levels(comb$cluster_prim)[levels(comb$cluster_prim) %in% unique(comb$cluster_prim)],]
met_cols = clin_meta_cols[cluster %in% levels(comb$cluster_met)[levels(comb$cluster_met) %in% unique(comb$cluster_met)],]
prim_cols = rev(prim_cols[order(match(cluster,levels(comb$cluster_prim)[levels(comb$cluster_prim) %in% unique(comb$cluster_prim)])),]$cols)
met_cols = rev(met_cols[order(match(cluster,levels(comb$cluster_met)[levels(comb$cluster_met) %in% unique(comb$cluster_met)])),]$cols)

#Plot flowdiagram
ggplot(comb,aes(y = Freq, axis1 = cluster_prim , axis3 = cluster_met)) +
  geom_flow(stat = "alluvium", aes(fill = cluster_prim)) +
  scale_x_discrete(limits = c("cluster_prim","cluster_met")) +
  scale_fill_manual(values = rev(prim_cols))+
  geom_stratum(fill = c(prim_cols,met_cols))+
  geom_text(stat = "stratum", infer.label = TRUE)


#How often does most abundant cluster match
maj_prim = maj_prim[PID %in% maj_met$PID]
maj_prim = maj_prim[order(match(PID,maj_met$PID))]
maj_met = maj_met[PID %in% maj_prim$PID]
maj_met = maj_met[order(match(PID,maj_prim$PID))]
maj_prim$met_cluster = maj_met$cluster

#Assign to group different or same
maj_prim$grp[maj_prim$cluster == maj_prim$met_cluster] = "same"
maj_prim$grp[is.na(maj_prim$grp)] = "different"

p = maj_prim[,nPID := .N, by = c("grp")]
p = unique(p[,c("grp","nPID")])
g1 = ggplot(p,aes(x = as.factor(1),y = nPID, fill = grp))+geom_bar(stat = "identity")+scale_fill_grey()+
  theme(aspect.ratio = 2)+
  ggtitle("Majority cell type")


#How often is the most abundant disseminated cluster present in the primary tumor
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
#Exclude shitty images and cells
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
met[,test := .N,by = "PID"]
#met = met[test >=100,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('clin_meta_name','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

m = apply(met[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(met[,-"PID"])[x]})

#Exclude ones with shared majority
PID_excl = met$PID[unlist(lapply(n,length)) > 1]
n = n[!unlist(lapply(n,length)) > 1]

maj_met = as.data.table(unlist(n))
names(maj_met) = "cluster"
maj_met$PID = met$PID[!met$PID %in% PID_excl]

#Primary
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
#Exclude shitty images and cells
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
primary[,test := .N,by = "PID"]
#primary = primary[test >=100,]
primary[,frac := ncells/.N,by = "PID"]

maj_met = maj_met[PID %in% primary$PID,]

#Check how many in primary for each
maj_met$fracPrim = NA
for (i in 1:nrow(maj_met)){
  curPID = maj_met$PID[i]
  curClus = maj_met$cluster[i]
  
  unIDs = unique(primary[PID == curPID & clin_meta_name == curClus]$id)
  maj_met$existsPrim[i] = length(unIDs)
  if (length(unIDs) > 0){
    maj_met$fracPrim[i] = unique(primary[PID == curPID & clin_meta_name == curClus]$frac)
  }
}

#At least 5 cells
maj_met$existsLog[maj_met$existsPrim < 5 ] = "absent"
maj_met$existsLog[is.na(maj_met$existsLog)] = "present"

p = maj_met[,nPID := .N, by = c("existsLog")]
p = unique(p[,c("existsLog","nPID")])
g2 = ggplot(p,aes(x = as.factor(1),y = nPID, fill = existsLog))+geom_bar(stat = "identity")+scale_fill_grey()+
  theme(aspect.ratio = 2)+
  ggtitle("At least 5 cells")

#plot next to each other
g1 + g2

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/clin_clust.pdf")
g1 + g2
dev.off()

```


Paired differential abundance analysis of simplified tumor clusters between matched primary and LN met samples
```{r differential abundance}
#Extract relevant information for primary tumor and LN met samples
both = as.data.table(data.frame(colData(subset(sce_t,,TissueType %in% c("primary tumor","lymph node mts")))[c('TissueType','PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType %in% c("primary tumor","lymph node mts")))))

#Exclude sizeflagged cells
both = both[sizeflag != 1,]

#Only use samples with matched primary and LN met tissue
both[,test:= length(unique(TissueType)),by = "PID"]
both = both[test == 2,]

#Calculate number of cells from each cluster in primary and LN met sample of each patient
both[,ncells:=.N ,by= c('PID','TissueType', 'clin_meta_name')]
both = unique(both[,c('clin_meta_name','ncells','PID','TissueType')])

#Prepare data for differential abundance analysis using edgeR
both = dcast.data.table(both, 'PID + TissueType ~ clin_meta_name',value.var = 'ncells',fill = 0)
both$TissueType = factor(both$TissueType,levels =c("primary tumor","lymph node mts"))
both$PID = factor(both$PID)

#Create design matrix
design <- createDesignMatrix(
  both[,c("PID","TissueType")], cols_design = c( "PID","TissueType")
)

#Create contrast
contrast <- createContrast(c(rep(0, ncol(design)-1),1))

#Check contrast
data.frame(parameters = colnames(design), contrast)

#Normalize
norm_factors <- calcNormFactors(t(both[,-c("PID","TissueType")]), method = "TMM")

#Differential abundance
y <- DGEList(t(both[,-c("PID","TissueType")]), norm.factors = norm_factors)
y <- estimateDisp(y,design)

fit <- glmFit(y, design)
lrt <- glmLRT(fit, contrast = contrast)
top <- edgeR::topTags(lrt, n = Inf, adjust.method = "BH", 
        sort.by = "none")

#Write out result table
res = as.data.table(top$table)
res$cluster = rownames(top$table)
res$sign = 0
res$sign[res$FDR < 0.01] = 1
res[order(FDR)]

#Plot logFC of significant clusters
res = res[sign == 1,]
res[logFC > 0,posneg := "pos"]
res[logFC < 0,posneg := "neg"]
res = res[order(logFC)]
res[,cluster := factor(cluster,levels = cluster)]

ggplot(res,aes(y = cluster, x = logFC, fill = posneg))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c("lightblue","purple"))

ggplot(res,aes(y = cluster, x = logFC, fill = cluster))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = clin_meta_cols[order(match(cluster,res$cluster))]$cols)

```

EMdist between primary tumor and matched LN met for all markers
```{r}
library(emdist)

#Get marker single-cell counts and relevant metadata for matched samples
dat = data.table(id = colnames(subset(sce_t,,sizeflag != 1)),t(counts(subset(sce_t,,sizeflag != 1))),PID = subset(sce_t,,sizeflag != 1)$PID, TissueType = subset(sce_t,,sizeflag != 1)$TissueType)
dat = dat[TissueType %in% c("primary tumor","lymph node mts"),]
dat[,ntissues := length(unique(TissueType)), by = c("PID")]
dat = dat[ntissues == 2,]
#Kick out images with few cells
dat[,ncells := .N,by = "PID"]
dat = dat[ncells>=100,]
dat$ncells = NULL

# #PR
# ggplot(cur, aes(`Gd158_Progesterone Receptor A/B`, fill=TissueType)) + geom_density(alpha=0.5)

#Prepare data.frame to save output into
pids = data.frame(PID = unique(dat$PID))
m = as.data.frame(matrix(0,nrow(pids),length(colnames(dat)[!colnames(dat) %in% c("TissueType","ntissues","id","PID")])))
colnames(m) = colnames(dat)[!colnames(dat) %in% c("TissueType","ntissues","id","PID")]
emds = cbind(pids,m)

#Loop through all patients
for (i in 1:length(unique(dat$PID))){
  cur = dat[PID == unique(dat$PID)[i]]
  
  #Loop through all markers
  for (marker in colnames(cur)[!colnames(cur) %in% c("PID","TissueType","ntissues","id")]){
    
      #Calculate emd between primary and met for current marker
      c = density(as.matrix(cur[TissueType == "primary tumor",eval(marker),with = F]))
      c_mat = as.matrix(data.frame(x = c$y,y = c$x))
      
      d = density(as.matrix(cur[TissueType == "lymph node mts",eval(marker),with = F]))
      d_mat = as.matrix(data.frame(x = d$y,y = d$x))
      
      emds[i,marker] = emd(c_mat,d_mat)
  }
}

emds = as.data.table(emds)
fwrite(emds,"/mnt/ZTMA_21_25_26/analysis/emds_primary_met.csv",col.names = T)

emd_long = melt.data.table(emds,id.vars = "PID",value.name = "emds",variable.name = "channel")


ggplot(emd_long,aes(x = channel,y =emds))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90))


#Compare HH3 between any 2 primary tumors or LN mets
set.seed(2)
emds_prim = rep(0,length(unique(dat$PID)))

#Loop through all patients so it results in same amount of comparisons as above
for (i in 1:length(unique(dat$PID))){
  rand = sample(unique(dat$PID),2, replace = F)
  cur = dat[PID %in% rand,]
    
  #Calculate emd between 2 different primaries for HH3
  c = density(as.matrix(cur[PID == rand[1],"In113_Histone H3",with = F]))
  c_mat = as.matrix(data.frame(x = c$y,y = c$x))
  
  d = density(as.matrix(cur[PID == rand[2],"In113_Histone H3",with = F]))
  d_mat = as.matrix(data.frame(x = d$y,y = d$x))
  
  emds_prim[i] = emd(c_mat,d_mat)

}


emds = fread("/mnt/ZTMA_21_25_26/analysis/emds_primary_met.csv",header = T)
comb = emds[,c("PID","In113_Histone H3")]
comb$type = "prim_met"

prim = data.table(PID = 1000,`In113_Histone H3` = emds_prim,type = "prim_prim")
comb = rbind(comb,prim)

ggplot(comb,aes(x =type ,y =`In113_Histone H3`))+geom_boxplot()+theme(axis.text.x = element_text(angle = 90))

```

Dot plot fractions in primary vs met but with clin_meta_name
```{r}
#Met
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
met[,test := .N,by = "PID"]
met = met[test >=100,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('clin_meta_name','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

#Primary
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
primary[,test := .N,by = "PID"]
primary = primary[test >=100,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('clin_meta_name','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

met = met[PID %in% primary$PID,]
primary = primary[PID %in% met$PID,]
met = met[order(match(PID, primary$PID)),]


met = melt.data.table(met,id.vars = "PID")
primary = melt.data.table(primary,id.vars = "PID")

comb = merge(met,primary,by = c("PID","variable"))
colnames(comb) = c("PID","cluster","frac_met","frac_primary")

#Exclude ones that are not present in primary or met
comb = comb[!(frac_met == 0 & frac_primary == 0),]

ggplot(comb,aes(x = frac_met, y = frac_primary,color = cluster))+
  geom_point()+
  scale_color_manual(values = clin_meta_cols[order(match(cluster,levels(comb$cluster))),]$col)

#Correlation
cor(comb$frac_met,comb$frac_primary)

```

Same analysis but across cores from the same primary tumor in Zurich dataset
```{r}
#Read data from previous analysis
Zurich_dat = fread("/mnt/ZTMA_21_25_26/analysis/ZurichTMA_clusters.csv",header = T)
Zurich_dat = Zurich_dat[!location == "[]"]

#Only tumor cells
Zurich_dat = Zurich_dat[cluster %in% c(14:27),]

#Calculate fractions per core
Zurich_dat[,ncells := .N,by = c("cluster","core")]
Zurich_dat[,frac := ncells/.N, by = "core"]
Zurich_dat = unique(Zurich_dat[,-c("id","ncells")])

#Select 2 from each patient
Zurich_dat[,ncores := length(unique(core)),by = "PID"]
Zurich_dat = Zurich_dat[ncores > 1,]
sample = lapply(unique(Zurich_dat$PID),function(x){sample(unique(Zurich_dat[PID == x,]$core),2,replace = F)})
Zurich_dat = Zurich_dat[core %in% unlist(sample),]
Zurich_dat[,ncores := length(unique(core)),by = "PID"] #check

#Select one to be analogous to primary and other to met
sample = lapply(unique(Zurich_dat$PID),function(x){sample(unique(Zurich_dat[PID == x,]$core),1,replace = F)})
one = Zurich_dat[core %in% sample,]
other = Zurich_dat[!core %in% sample,]
setnames(one, "frac", "frac_one")
setnames(other, "frac", "frac_other")

comb = merge(one[,c("core","PID","frac_one","cluster")],other[,c("core","PID","frac_other","cluster")],by = c("PID","cluster"))
comb$cluster = as.factor(comb$cluster)

#Metacluster colors from previous paper
mycols_basel_meta <- colors()[c(258,257,86,85,259,81, #green
                                652, #yellow
                                636,520,430,109,68,43,#light blue
                                132,#other dark blue
                                26,#dark blue
                                8,12, #turquouis
                                33,#red
                                551,#dark purple
                                95,#light purple
                                419,#light pink
                                117,#pink
                                52,#burnt orange
                                500,#orange
                                568,#pink orange
                                624, #brown
                                133)] #dark red

#Exclude ones that are not present in primary or met
comb = comb[!(frac_one == 0 & frac_other == 0),]

ggplot(comb,aes(x = frac_one, y = frac_other,color = cluster))+
  geom_point()+
  scale_color_manual(values = mycols_basel_meta[14:27])

#Correlation
cor(comb$frac_one,comb$frac_other)
```

KL divergence across cores from the same primary tumor in Zurich dataset
```{r}
library(entropy)

#Read data from previous analysis
Zurich_dat = fread("/mnt/ZTMA_21_25_26/analysis/ZurichTMA_clusters.csv",header = T)
Zurich_dat = Zurich_dat[!location == "[]"]

#Only tumor cells
Zurich_dat = Zurich_dat[cluster %in% c(14:27),]

#Calculate fractions per core
Zurich_dat[,ncells := .N,by = c("cluster","core")]
Zurich_dat[,frac := ncells/.N, by = "core"]
Zurich_dat = unique(Zurich_dat[,-c("id","ncells")])

#Fill in missing ones with almost 0 (Dividing by zero in KL divergence results in inf)
Zurich_dat =dcast.data.table(Zurich_dat,"core + PID + location ~ cluster",value.var = "frac",fill = 0.000001)
Zurich_dat = melt.data.table(Zurich_dat,id.vars = c("core","PID","location"),variable.name = "cluster",value.name = "frac")

#Select 2 from each patient
Zurich_dat[,ncores := length(unique(core)),by = "PID"]
Zurich_dat = Zurich_dat[ncores > 1,]
sample = lapply(unique(Zurich_dat$PID),function(x){sample(unique(Zurich_dat[PID == x,]$core),2,replace = F)})
Zurich_dat = Zurich_dat[core %in% unlist(sample),]
Zurich_dat[,ncores := length(unique(core)),by = "PID"] #check

#Select one to be analogous to primary and other to met
sample = lapply(unique(Zurich_dat$PID),function(x){sample(unique(Zurich_dat[PID == x,]$core),1,replace = F)})
one = Zurich_dat[core %in% sample,]
other = Zurich_dat[!core %in% sample,]
setnames(one, "frac", "frac_one")
setnames(other, "frac", "frac_other")

KL_zurich = merge(one[,c("core","PID","frac_one","cluster")],other[,c("core","PID","frac_other","cluster")],by = c("PID","cluster"))
KL_zurich$cluster = as.factor(KL_zurich$cluster)

#KL divergence
KL_zurich[,KLdiv := KL.plugin(frac_one,frac_other),by = "PID"]

```

KL divergence between primary and met on clinical clustering
```{r}
#Met
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
met[,test := .N,by = "PID"]
met = met[test >=100,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('clin_meta_name','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

#Primary
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','clin_meta_name','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'clin_meta_name')]
primary[,test := .N,by = "PID"]
primary = primary[test >=100,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('clin_meta_name','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ clin_meta_name',value.var = 'frac_cluster',fill = 0)

met = met[PID %in% primary$PID,]
primary = primary[PID %in% met$PID,]
met = met[order(match(PID, primary$PID)),]


met = melt.data.table(met,id.vars = "PID")
primary = melt.data.table(primary,id.vars = "PID")

KL_clin = merge(met,primary,by = c("PID","variable"))
colnames(KL_clin) = c("PID","cluster","frac_met","frac_primary")

#Add small number to zeros so KL divergence doesn't become inf
KL_clin[frac_met == 0,frac_met := 0.000001]
KL_clin[frac_primary == 0,frac_primary := 0.000001]

#KL divergence
KL_clin[,KLdiv := KL.plugin(frac_primary,frac_met),by = "PID"]

```

KL div on 59 tumor clusters
```{r}
#Read in tumor sce
sce_t_orig = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE_T.rds')


#Met
met = as.data.table(data.frame(colData(subset(sce_t_orig,,TissueType == "lymph node mts"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t_orig,,TissueType == "lymph node mts"))))
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
met[,test := .N,by = "PID"]
met = met[test >=100,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('tumor_cluster_100','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

#Primary
primary = as.data.table(data.frame(colData(subset(sce_t_orig,,TissueType == "primary tumor"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t_orig,,TissueType == "primary tumor"))))
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
primary[,test := .N,by = "PID"]
primary = primary[test >=100,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('tumor_cluster_100','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

met = met[PID %in% primary$PID,]
primary = primary[PID %in% met$PID,]
met = met[order(match(PID, primary$PID)),]


met = melt.data.table(met,id.vars = "PID")
primary = melt.data.table(primary,id.vars = "PID")

KL_59 = merge(met,primary,by = c("PID","variable"))
colnames(KL_59) = c("PID","cluster","frac_met","frac_primary")

#Add small number to zeros so KL divergence doesn't become inf
KL_59[frac_met == 0,frac_met := 0.000001]
KL_59[frac_primary == 0,frac_primary := 0.000001]

#KL divergence
KL_59[,KLdiv := KL.plugin(frac_primary,frac_met),by = "PID"]


```

Compare KL divs as boxplots
```{r}
KL_clin$type = "clinical clustering prim-met"
KL_zurich$type = "Zurich"
KL_59$type = "59 clusters"

comb = rbind(unique(KL_zurich[,c("PID","KLdiv","type")]),unique(KL_clin[,c("PID","KLdiv","type")]),unique(KL_59[,c("PID","KLdiv","type")]))

ggplot(comb, aes(x = type,y = KLdiv))+ geom_boxplot()

```

Compare tumor cells without panCK with stromal cells
```{r}
sce_t = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE_T.rds')
sce_s = readRDS('/mnt/ZTMA_21_25_26/analysis/SCE_S.rds')
tcols = fread('/mnt/ZTMA_21_25_26/analysis/tcols.csv',header = T)

#Tumor colors
t = data.table(id = colnames(sce_t),panCK = t(counts(sce_t["Lu175_Keratin Epithelial",])),type = sce_t$tumor_cluster_100)
s = data.table(id = colnames(sce_s),panCK = t(counts(sce_s["Lu175_Keratin Epithelial",])),type = "stroma")
comb = rbind(t,s)

ggplot(comb,aes(x = `panCK.Lu175_Keratin Epithelial`,color = type))+geom_density() + scale_color_manual(values = c(tcols$tumor_cols,"black"))#,y=..scaled..

#Only 58 vs stroma
comb = comb[type %in% c("stroma","58"),]
comb$type = factor(comb$type,levels = c("stroma","58"))
ggplot(comb,aes(x = `panCK.Lu175_Keratin Epithelial`,fill = type,alpha = 0.5))+geom_density() + scale_fill_manual(values = c("black",tcols[cluster == 58,]$tumor_cols))

#All tumor vs all stroma
t = data.table(id = colnames(sce_t),panCK = t(counts(sce_t["Lu175_Keratin Epithelial",])),type = "tumor")
s = data.table(id = colnames(sce_s),panCK = t(counts(sce_s["Lu175_Keratin Epithelial",])),type = "stroma")
comb = rbind(t,s)

#comb = comb[type == "tumor",]
ggplot(comb,aes(x = `panCK.Lu175_Keratin Epithelial`,fill = type,alpha = 0.5))+geom_density()#+ylim(0,50)
  

```

Find patients for which we need whole sections
```{r}
#Get most abundant LN met cluster
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
met[,test := .N,by = "PID"]
met = met[test >=20,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('tumor_cluster_100','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

#Extract for each sample the cluster that makes up the largest proportion of cells
m = apply(met[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(met[,-"PID"])[x]})

#Exclude samples with shared largest proportion
PID_excl = met$PID[unlist(lapply(n,length)) > 1]
n = n[!unlist(lapply(n,length)) > 1]

#Put most abundant LN met cluster for each patient into a table
maj_met = as.data.table(unlist(n))
names(maj_met) = "cluster_met"
maj_met$PID = met$PID[!met$PID %in% PID_excl]


#Get most abundant primary tumor cluster
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
primary[,test := .N,by = "PID"]
primary = primary[test >=20,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('tumor_cluster_100','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

#Extract for each sample the cluster that makes up the largest proportion of cells
m = apply(primary[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(primary[,-"PID"])[x]})

#Exclude samples with shared largest proportion
PID_excl = primary$PID[unlist(lapply(n,length)) > 1]
n = n[!unlist(lapply(n,length)) > 1]

#Put most abundant primary tumor cluster for each patient into a table
maj_prim = as.data.table(unlist(n))
names(maj_prim) = "cluster_prim"
maj_prim$PID = primary$PID[!primary$PID %in% PID_excl]


#Combine primary and LN met 
comb = merge(maj_prim,maj_met,by = "PID")

#add clinical classification
add = unique(as.data.table(colData(subset(sce_t,,TissueType == "primary tumor"))[,c("PID","Mol_Signature","core","ER_pr")]))
add[duplicated(PID),Mol_Signature := "#NULL!"]
add[duplicated(PID,fromLast = T),Mol_Signature := "#NULL!"]
add = unique(add)
comb = merge(comb,add,by = "PID",all.x = T)
setnames(comb,"core","core_primary")

#add met cores
add = unique(as.data.table(colData(subset(sce_t,,TissueType == "lymph node mts"))[,c("PID","core")]))
comb = merge(comb,add,by = "PID",all.x = T)
setnames(comb,"core","core_met")

comb[cluster_met == "27",]
comb[cluster_met == "7",]
comb[cluster_met == "29",]

#selection
PIDs_selected_27 = c("569","713","126","761")
PIDs_selected_7 = c("99","402","488","969")
PIDs_selected_29 = c("171","1364")

all = c(PIDs_selected_27,PIDs_selected_7,PIDs_selected_29)

```

Find node-negative primary tumors with dendritic cells (correlation is slightly better with numbers of cells instead of fractions -> maybe use density)
```{r}
#node neg primary tumors with cluster 29
d = as.data.table(colData(subset(sce_t,,TissueType == "primary tumor" & pN_4gr == "pN0" & sizeflag != 1))[,c("PID","tumor_cluster_100")])
d[,test := .N,by = "PID"]
d = d[test >= 100,]
d[,ncells := .N ,by = c("PID","tumor_cluster_100")]
d[,frac_29 := ncells/.N,by = "PID"]
d = unique(d[tumor_cluster_100 == "29",])

#node neg primary tumors stromal clusters
s = as.data.table(colData(subset(sce_s,,TissueType == "primary tumor" & pN_4gr == "pN0" & sizeflag != 1))[,c("PID","stroma_cluster")])
s[,test := .N,by = "PID"]
s = s[test >= 100,]
s[,ncells := .N ,by = c("PID","stroma_cluster")]
s[,frac := ncells/.N,by = "PID"]
s = unique(s)
s = dcast.data.table(s,"PID ~ stroma_cluster",value.var = "frac",fill = 0)

d = merge(s,d[,c("PID","frac_29")],by = "PID",all.x = T)
d[is.na(frac_29),frac_29:=0]

c = cor(as.matrix(d[,-"PID"]))

tcols = fread('/mnt/ZTMA_21_25_26/analysis/tcols.csv',header = T)
scols = fread('/mnt/ZTMA_21_25_26/analysis/scols.csv',header = T)
setnames(scols,"stroma_cols","cols")
setnames(tcols,"tumor_cols","cols")
tcols[,cluster := as.character(cluster)]
tcols[cluster == 29,cluster := "frac_29"]
cols = rbind(scols,tcols[cluster == "frac_29",])

#clustered heatmap
ha = HeatmapAnnotation(tumor = colnames(c),
    col = list(tumor = structure(cols$cols,names = cols$cluster)))
hm = Heatmap(c, name = "correlation",column_title = "stroma primary", km = 1, col = colorRamp2(c(-1,0, 1), c('mediumpurple3',"white", "darkorange")),
            show_row_names = T, show_column_names =  T, clustering_method_rows = "ward.D2",clustering_method_columns = "ward.D2",top_annotation = ha)+
  rowAnnotation(rn = anno_text(rownames(c)))

cor(d$frac_29,s$`8`)

ggplot(d,aes(x = frac_29,y = `8`))+geom_point()

cores = unique(as.data.table(colData(subset(sce_t,,TissueType == "primary tumor"))[,c("PID","core")]))
d = merge(d,cores,by = "PID")
d[frac_29 > 0.05 & `8` > 0.05]
d[frac_29 > 100 & `8`> 50] #when using numbers of cells instead of fractions

#Get whole slides for dendritic cells
sel_neg = c("256","668","1771","1545","1415")


#Are there generally more dendritic cells in primary tumors?
s = as.data.table(colData(subset(sce_s,,TissueType == "primary tumor" & sizeflag != 1))[,c("PID","stroma_cluster","pN_4gr")])
s[,test := .N,by = "PID"]
s = s[test >= 100,]
s[,ncells := .N ,by = c("PID","stroma_cluster")]
s[,frac := ncells/.N,by = "PID"]
s = unique(s[stroma_cluster == "8",])

ggplot(s,aes(x = pN_4gr, y = frac))+geom_boxplot()


#Check for node-positive tumors with cluster 29
d = as.data.table(colData(subset(sce_t,,TissueType == "primary tumor" & pN_4gr %in% c("pN1","pN2","pN3") & sizeflag != 1))[,c("PID","tumor_cluster_100")])
d[,test := .N,by = "PID"]
d = d[test >= 100,]
d[,ncells := .N ,by = c("PID","tumor_cluster_100")]
d[,frac_29 := ncells/.N,by = "PID"]
d = unique(d[tumor_cluster_100 == "29",])


s = as.data.table(colData(subset(sce_s,,TissueType == "primary tumor" & pN_4gr %in% c("pN1","pN2","pN3") & sizeflag != 1))[,c("PID","stroma_cluster")])
s[,test := .N,by = "PID"]
s = s[test >= 100,]
s[,ncells := .N ,by = c("PID","stroma_cluster")]
s[,frac := ncells/.N,by = "PID"]
s = unique(s)
s = dcast.data.table(s,"PID ~ stroma_cluster",value.var = "frac",fill = 0)
s = s[,c("PID","8")]

d = merge(s,d,by = "PID",all.x = T)
d[is.na(frac_29),frac_29 := 0]

cor(d$frac_29,d$`8`)

ggplot(d,aes(x = frac_29,y = `8`))+geom_point()

d = merge(d,cores,by = "PID")
d[frac_29 > 0.2 & `8` < 0.1]

sel = c("234","420","424","1433","1632")

#Or should I only use tumors where this cell type has acctually disseminated?
m = as.data.table(colData(subset(sce_t,,TissueType == "lymph node mts" & sizeflag != 1))[,c("PID","tumor_cluster_100")])
m[,test := .N,by = "PID"]
m = m[test >= 100,]
m[,ncells := .N ,by = c("PID","tumor_cluster_100")]
m[,frac_29 := ncells/.N,by = "PID"]
m = unique(m[tumor_cluster_100 == "29",])
m = m[ncells > 5,]
m[PID %in% sel,]


d = d[PID %in% m$PID,]
ggplot(d,aes(x = frac_29,y = `8`))+geom_point()
d[frac_29 > 0.2 & `8` < 0.05]

all = c(all,sel_neg,sel)
all = as.numeric(all)
sort(all)

#Check that TMAs are right
f = unique(as.data.table(colData(subset(sce_t,,TissueType == "primary tumor"))[,c("PID","core")]))
f = f[PID %in% all,]
f[order(as.numeric(as.character(PID))),]

```

#Is clinical classification phenotype always present (at least) as subpopulation in the primary
```{r}
#Primary
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','sizeflag','Mol_Signature')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
primary = cbind(primary, as.data.frame(t(counts(subset(sce_t,,TissueType == "primary tumor")))))
primary = primary[sizeflag != 1,]

ggplot(primary, aes(x = `Gd156_Estrogen Receptor Alpha`))+geom_density()

res <- densityMclust(primary$`Gd156_Estrogen Receptor Alpha`, 2)
primary$GMM_ER = res$classification
res <- densityMclust(primary$`Gd158_Progesterone Receptor A/B`, 2)
primary$GMM_PR = res$classification
res <- densityMclust(primary$`Eu151_c-erbB-2 - Her2`, 2)
primary$GMM_HER2 = res$classification
primary[`Eu151_c-erbB-2 - Her2`<0.5,GMM_HER2 :=1]

ggplot(primary, aes(x = `Eu151_c-erbB-2 - Her2`,fill = as.factor(GMM_HER2)))+geom_density()

primary[,test := .N,by = "PID"]
primary = primary[test >=20,]
primary[,ncells :=sum(GMM_ER == 2) ,by= 'PID']
primary[,frac_pos_ER := ncells/.N, by=c('PID')]

primary[,ncells :=sum(GMM_PR == 2) ,by= 'PID']
primary[,frac_pos_PR := ncells/.N, by=c('PID')]

primary[,ncells :=sum(GMM_HER2 == 2) ,by= 'PID']
primary[,frac_pos_HER2 := ncells/.N, by=c('PID')]

primary = unique(primary[,c("PID","frac_pos_ER","frac_pos_PR","frac_pos_HER2",'Mol_Signature')])
primary$HRpos_prim = 0
primary$HER2pos_prim = 0
primary[frac_pos_ER > 0.05 | frac_pos_PR > 0.05, HRpos_prim := 1]#It would actually be 0.05 but only claiming they are present not in at least 5%
primary[frac_pos_HER2 > 0.1 , HER2pos_prim := 1]
primary[HRpos_prim == 1 & HER2pos_prim == 1,ASCO := "HR+HER2+"]
primary[HRpos_prim == 1 & HER2pos_prim == 0,ASCO := "HR+HER2-"]
primary[HRpos_prim == 0 & HER2pos_prim == 1,ASCO := "HR-HER2+"]
primary[HRpos_prim == 0 & HER2pos_prim == 0,ASCO := "HR-HER2-"]

test = primary[Mol_Signature == "Luminal B (HER2-)",]

p = test[,nHER2 := .N, by = c("HRpos_prim")]
p = unique(p[,c("HRpos_prim","nHER2")])
g1 = ggplot(p,aes(x = as.factor(1),y = nHER2, fill = as.factor(HRpos_prim)))+geom_bar(stat = "identity")+scale_fill_grey()+
  theme(aspect.ratio = 2)+
  ggtitle("Luminal B (HER2-)")

test = primary[Mol_Signature == "Luminal A",]

p = test[,nHER2 := .N, by = c("HRpos_prim")]
p = unique(p[,c("HRpos_prim","nHER2")])
g2 = ggplot(p,aes(x = as.factor(1),y = nHER2, fill = as.factor(HRpos_prim)))+geom_bar(stat = "identity")+scale_fill_grey()+
  theme(aspect.ratio = 2)+
  ggtitle("Luminal A")

test = primary[Mol_Signature == "Luminal B (HER2+)",]

p = test[,nHER2 := .N, by = c("HER2pos_prim")]
p = unique(p[,c("HER2pos_prim","nHER2")])
g3 = ggplot(p,aes(x = as.factor(1),y = nHER2, fill = as.factor(HER2pos_prim)))+geom_bar(stat = "identity")+scale_fill_grey()+
  theme(aspect.ratio = 2)+
  ggtitle("Luminal B (HER2+)")

test = primary[Mol_Signature == "Triple negative",]

g1 + g2 + g3



#Flowdiagram (mol_sign -> primary ASCO/CAP -> LN met)
primary[,freq := .N, by = c("Mol_Signature","ASCO")]
primary[,Mol_Signature:= factor(Mol_Signature,levels = c("Luminal B (HER2-)","Luminal B (HER2+)","#NULL!","Triple negative","Luminal A","HER2"))]
test = unique(primary[,c("Mol_Signature","ASCO","freq")])

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/flow.pdf")
ggplot(test,aes(y = freq, axis1 = Mol_Signature , axis3 = ASCO)) +
  geom_flow(stat = "alluvium", aes(fill = Mol_Signature)) +
  scale_x_discrete(limits = c("Mol_Signature","ASCO")) +
  scale_fill_manual(values = cols_mol$col)+
  geom_stratum(fill = c(rev(cols_mol$col[1:6]),c("grey20","grey70","grey50","grey90")))+
  geom_text(stat = "stratum", infer.label = TRUE)
dev.off()


#Add primary and LN met

all = as.data.table(data.frame(colData(subset(sce_t,,TissueType %in% c("primary tumor","lymph node mts")))[c('PID','sizeflag','Mol_Signature',"TissueType")],id = colnames(subset(sce_t,,TissueType %in% c("primary tumor","lymph node mts")))))
all = cbind(all, as.data.frame(t(counts(subset(sce_t,,TissueType %in% c("primary tumor","lymph node mts"))))))
all = all[sizeflag != 1,]

ggplot(all, aes(x = `Gd156_Estrogen Receptor Alpha`))+geom_density()

res <- densityMclust(all$`Gd156_Estrogen Receptor Alpha`, 2)
all$GMM_ER = res$classification
res <- densityMclust(all$`Gd158_Progesterone Receptor A/B`, 2)
all$GMM_PR = res$classification
res <- densityMclust(all$`Eu151_c-erbB-2 - Her2`, 2)
all$GMM_HER2 = res$classification
all[`Eu151_c-erbB-2 - Her2`<0.5,GMM_HER2 :=1]

ggplot(all, aes(x = `Eu151_c-erbB-2 - Her2`,fill = as.factor(GMM_HER2)))+geom_density()


all[,test := .N,by = c("PID","TissueType")]
all = all[test >=20,]
all[,ncells :=sum(GMM_ER == 2) ,by= c("PID","TissueType")]
all[,frac_pos_ER := ncells/.N, by=c("PID","TissueType")]

all[,ncells :=sum(GMM_PR == 2) ,by= c("PID","TissueType")]
all[,frac_pos_PR := ncells/.N, by=c("PID","TissueType")]

all[,ncells :=sum(GMM_HER2 == 2) ,by= c("PID","TissueType")]
all[,frac_pos_HER2 := ncells/.N, by=c("PID","TissueType")]

all = unique(all[,c("PID","frac_pos_ER","frac_pos_PR","frac_pos_HER2",'Mol_Signature',"TissueType")])
all$HRpos_prim = 0
all$HER2pos_prim = 0
all[frac_pos_ER > 0.05 | frac_pos_PR > 0.05, HRpos_prim := 1]#It would actually be 0.05 but only claiming they are present not in at least 5%
all[frac_pos_HER2 > 0.1 , HER2pos_prim := 1]
all[HRpos_prim == 1 & HER2pos_prim == 1,ASCO := "HR+HER2+"]
all[HRpos_prim == 1 & HER2pos_prim == 0,ASCO := "HR+HER2-"]
all[HRpos_prim == 0 & HER2pos_prim == 1,ASCO := "HR-HER2+"]
all[HRpos_prim == 0 & HER2pos_prim == 0,ASCO := "HR-HER2-"]

dall = merge(all[TissueType == "primary tumor",],all[TissueType == "lymph node mts",c("PID","ASCO")],by = "PID",all.x = T)

#Flowdiagram (mol_sign -> primary ASCO/CAP -> LN met)
dall[,freq := .N, by = c("Mol_Signature","ASCO.x","ASCO.y")]
dall[,Mol_Signature:= factor(Mol_Signature,levels = c("Luminal B (HER2-)","Luminal B (HER2+)","#NULL!","Triple negative","Luminal A","HER2"))]
test = unique(dall[,c("Mol_Signature","ASCO.x","ASCO.y","freq")])

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/flow_pm.pdf")
ggplot(test,aes(y = freq, axis1 = Mol_Signature , axis2 = ASCO.x, axis3 = ASCO.y)) +
  geom_flow(stat = "alluvium", aes(fill = Mol_Signature)) +
  scale_x_discrete(limits = c("Mol_Signature","ASCO.x","ASCO.y")) +
  scale_fill_manual(values = cols_mol$col)+
  geom_stratum(fill = c(rev(cols_mol$col[1:6]),c("grey20","grey70","grey50","grey90"),c("white","grey20","grey70","grey50","grey90")))+
  geom_text(stat = "stratum", infer.label = TRUE)
dev.off()



#Only matched
dall = merge(all[TissueType == "primary tumor",],all[TissueType == "lymph node mts",c("PID","ASCO")],by = "PID")
dall[PID == "175",Mol_Signature := "HER2"]
dall[PID == "1569",Mol_Signature := "HER2"]
dall[PID == "552",Mol_Signature := "Luminal B (HER2+)"]
dall[PID == "1590",Mol_Signature := "Luminal A"]
dall = unique(dall)

#Flowdiagram (mol_sign -> primary ASCO/CAP -> LN met)
dall[,freq := .N, by = c("Mol_Signature","ASCO.x","ASCO.y")]
dall[,Mol_Signature:= factor(Mol_Signature,levels = c("Luminal B (HER2-)","Luminal B (HER2+)","#NULL!","Triple negative","Luminal A","HER2"))]
test = unique(dall[,c("Mol_Signature","ASCO.x","ASCO.y","freq")])

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/flow_pm.pdf")
ggplot(test,aes(y = freq, axis1 = Mol_Signature , axis2 = ASCO.x, axis3 = ASCO.y)) +
  geom_flow(stat = "alluvium", aes(fill = Mol_Signature)) +
  scale_x_discrete(limits = c("Mol_Signature","ASCO.x","ASCO.y")) +
  scale_fill_manual(values = cols_mol$col)+
  geom_stratum(fill = c(rev(cols_mol$col[1:6]),c("grey20","grey70","grey50","grey90"),c("grey20","grey70","grey50","grey90")))+
  geom_text(stat = "stratum", infer.label = TRUE)
dev.off()

#Withou clin subtype
dall[,freq := .N, by = c("ASCO.x","ASCO.y")]
test = unique(dall[,c("ASCO.x","ASCO.y","freq")])
test$ASCO.x = factor(test$ASCO.x,levels = rev(c("HR-HER2-","HR+HER2-","HR+HER2+","HR-HER2+")))
test$ASCO.y = factor(test$ASCO.y,levels = rev(c("HR-HER2-","HR+HER2-","HR+HER2+","HR-HER2+")))

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/HER2_push_flow.pdf")
ggplot(test,aes(y = freq, axis2 = ASCO.x, axis3 = ASCO.y)) +
  geom_flow(stat = "alluvium", aes(fill = ASCO.x)) +
  scale_x_discrete(limits = c("ASCO.x","ASCO.y")) +
  scale_fill_manual(values = rev(c("grey90","grey70","grey50","grey20")))+
  geom_stratum(fill = c(c("grey90","grey70","grey50","grey20"),c("grey90","grey70","grey50","grey20")))+
  geom_text(stat = "stratum", infer.label = TRUE)
dev.off()


#How many agree based on ASCO/CAP
m = all[TissueType == "lymph node mts",c("PID","ASCO","HRpos_prim","HER2pos_prim")]
colnames(m) = c("PID","ASCO_met","HRpos_met","HER2pos_met")
dall = merge(all[TissueType == "primary tumor",],m,by = "PID")
dall[duplicated(PID),]
dall[duplicated(PID,fromLast = T),]
dall[PID == "175",Mol_Signature := "HER2"]
dall[PID == "1569",Mol_Signature := "HER2"]
dall[PID == "552",Mol_Signature := "Luminal B (HER2+)"]
dall[PID == "1590",Mol_Signature := "Luminal A"]
dall = unique(dall)

dall[HRpos_prim == HRpos_met, HR := "agree"]
dall[HRpos_prim < HRpos_met, HR := "HR not in primary"]
dall[HRpos_prim > HRpos_met, HR := "HR not in met"]
dall[HER2pos_prim == HER2pos_met, HER2 := "agree"]
dall[HER2pos_prim < HER2pos_met, HER2 := "HER2 not in primary"] 
dall[HER2pos_prim > HER2pos_met, HER2 := "HER2 not in met"] 

p = dall[,nHR := .N, by = c("HR")] 
p = unique(p[,c("HR","nHR")])
g1 = ggplot(p,aes(x = as.factor(1),y = nHR, fill = as.factor(HR)))+geom_bar(stat = "identity")+scale_fill_grey()+ theme(aspect.ratio = 2)+ 
ggtitle("HR agree")

p = dall[,nHER2 := .N, by = c("HER2")]
 p = unique(p[,c("HER2","nHER2")])
g2 = ggplot(p,aes(x = as.factor(1),y = nHER2, fill = as.factor(HER2)))+geom_bar(stat = "identity")+scale_fill_grey()+ 
 theme(aspect.ratio = 2)+ 
ggtitle("HER2 agree")
pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/prop.pdf")
g1 + g2
dev.off()

#Agreement betw. primary ASCO and clinical subtype
dall$clinHR = "disagree"
dall[HRpos_prim == 1 & str_detect(Mol_Signature,"Luminal"), clinHR := "agree"]
dall[HRpos_prim == 0 & !str_detect(Mol_Signature,"Luminal"), clinHR := "agree"]
dall[Mol_Signature == "#NULL!",clinHR := "agree"]
dall[clinHR == "disagree" & HRpos_prim == HRpos_met,clinHR := "disagree BUT"]
dall$clinHR = factor(dall$clinHR, levels = c("disagree","disagree BUT","agree"))

dall$clinHER2 = "disagree"
dall[HER2pos_prim == 1 & Mol_Signature %in% c("Luminal B (HER2+)","HER2"), clinHER2 := "agree"]
dall[HER2pos_prim == 0 & !Mol_Signature %in% c("Luminal B (HER2+)","HER2"), clinHER2 := "agree"]
dall[Mol_Signature == "#NULL!",clinHER2 := "agree"]
dall[clinHER2 == "disagree"]
dall[clinHER2 == "disagree" & HER2pos_prim == HER2pos_met,clinHER2 := "disagree BUT"]
dall$clinHER2 = factor(dall$clinHER2, levels = c("disagree","disagree BUT","agree"))

p = dall[,nHR := .N, by = c("clinHR")] 
p = unique(p[,c("clinHR","nHR")])
g1 = ggplot(p,aes(x = as.factor(1),y = nHR, fill = as.factor(clinHR)))+geom_bar(stat = "identity")+scale_fill_grey(start=0.8, end=0.2)+ theme(aspect.ratio = 2)+ 
ggtitle("HR clin agree")

p = dall[,nHER2 := .N, by = c("clinHER2")]
 p = unique(p[,c("clinHER2","nHER2")])
g2 = ggplot(p,aes(x = as.factor(1),y = nHER2, fill = as.factor(clinHER2)))+geom_bar(stat = "identity")+scale_fill_grey(start=0.8, end=0.2)+ 
 theme(aspect.ratio = 2)+ 
ggtitle("HER2 clin agree")

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/HER2_push_clin_bars.pdf")
g1 + g2
dev.off()

```

Force regular flowdiagram into same order.....
```{r}
#Get most abundant LN met cluster
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
met[,test := .N,by = "PID"]
met = met[test >=20,] #met = met[test >=20,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('tumor_cluster_100','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

#Extract for each sample the cluster that makes up the largest proportion of cells
m = apply(met[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(met[,-"PID"])[x]})

#Exclude samples with shared largest proportion -> don't exclude just randomly sample!!!!
# PID_excl = met$PID[unlist(lapply(n,length)) > 1]
# n = n[!unlist(lapply(n,length)) > 1]

n[unlist(lapply(n,length)) > 1] = lapply(n[unlist(lapply(n,length)) > 1],function(x){x[[1]][1]})

#Put most abundant LN met cluster for each patient into a table
maj_met = as.data.table(unlist(n))
names(maj_met) = "cluster_met"
maj_met$PID = met$PID #[!met$PID %in% PID_excl]


#Get most abundant primary tumor cluster
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
primary[,test := .N,by = "PID"]
primary = primary[test >=20,] #primary = primary[test >=20,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('tumor_cluster_100','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

#Extract for each sample the cluster that makes up the largest proportion of cells
m = apply(primary[,-"PID"],1,function(x){x == max(x)})
n = apply(t(m),1,which)
n = lapply(n,function(x){names(primary[,-"PID"])[x]})

#Exclude samples with shared largest proportion -> nooo just randomly sample!
# PID_excl = primary$PID[unlist(lapply(n,length)) > 1]
# n = n[!unlist(lapply(n,length)) > 1]
n[unlist(lapply(n,length)) > 1] = lapply(n[unlist(lapply(n,length)) > 1],function(x){x[[1]][1]})


#Put most abundant primary tumor cluster for each patient into a table
maj_prim = as.data.table(unlist(n))
names(maj_prim) = "cluster_prim"
maj_prim$PID = primary$PID #[!primary$PID %in% PID_excl]


#Combine primary and LN met info and calculate shared patient fraction per most abundant cluster combination
comb = merge(maj_prim,maj_met,by = "PID")
add = unique(as.data.table(colData(subset(sce_t,,TissueType == "primary tumor"))[,c("PID","Mol_Signature")]))
add[PID == "175",Mol_Signature := "HER2"]
add[PID == "1569",Mol_Signature := "HER2"]
add[PID == "552",Mol_Signature := "Luminal B (HER2+)"]
add[PID == "1590",Mol_Signature := "Luminal A"]
add[duplicated(PID),Mol_Signature := "#NULL!"]
add[duplicated(PID,fromLast = T),Mol_Signature := "#NULL!"]
add = unique(add)

comb = merge(comb,add,by = "PID",all.x = T)

#order according to dall from ggalluvial above
comb = merge(comb,dall[,c("PID","ASCO.x")],by = "PID")
comb = comb[order(ASCO.x,cluster_prim),]
order_conceptually = rev(c("42","33","10","43","29","19","5","55","41","37","44","45","15","48","6","18","31","50","36","25","58","39","52","54","57","59","7","38","22","11","4","2","27","46","32","34","23","20","3","49","8","13","14","24","28","21","53","9","35","47","16","12","17","30","51","1","26"))
order_conceptually = order_conceptually[order_conceptually %in% unique(c(unique(comb$cluster_prim),unique(comb$cluster_met)))]
comb[,cluster_prim := factor(cluster_prim,levels = order_conceptually)]
comb[,cluster_met := factor(cluster_met,levels = order_conceptually)]

comb[,Freq := .N,by = c("cluster_prim","cluster_met")]
comb = unique(comb[,c("cluster_prim","cluster_met","Freq")])

prim_cols = tcols[cluster %in% levels(comb$cluster_prim)[levels(comb$cluster_prim) %in% unique(comb$cluster_prim)],]
prim_cols = prim_cols[order(match(cluster,levels(comb$cluster_prim)[levels(comb$cluster_prim) %in% unique(comb$cluster_prim)])),]$tumor_cols
met_cols = tcols[cluster %in% levels(comb$cluster_met)[levels(comb$cluster_met) %in% unique(comb$cluster_met)],]
met_cols = met_cols[order(match(cluster,levels(comb$cluster_met)[levels(comb$cluster_met) %in% unique(comb$cluster_met)])),]$tumor_cols


#Plot flow diagram
g = ggplot(comb,aes(y = Freq,axis1 = cluster_prim , axis2 = cluster_met)) +
  geom_flow(aes(fill = cluster_prim)) +
  scale_x_discrete(limits = c("cluster_prim","cluster_met")) +
  scale_fill_manual(values = prim_cols)+
  geom_stratum(fill = c(rev(prim_cols),rev(met_cols)))+
  geom_text(stat = "stratum", infer.label = TRUE)

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/flow2.pdf")
g
dev.off()

```

IF analysis triple-negative questions panCK = Cy3 and clin = Cy7
```{r IF analysis, eval = F}
dat1 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/Tripleneg_question/Segmentation/panCK_clin_first5/measurements_first5.csv")
dat2 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/Tripleneg_question/Segmentation/panCK_clin_second3/measurements_second3.csv")
dat3 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/Tripleneg_question/Segmentation/panCK_clin_second_last1/measurements_secondlast1.tsv")
dat4 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/Tripleneg_question/Segmentation/panCK_clin_last1/measurements_last1.csv")

#Combine
dat = rbind(dat1,dat2,dat3,dat4)

#Save out IF single-cll data
fwrite(dat,"/mnt/ZTMA_21_25_26/IF_singlecell/TN_question/panCK_clin_singlecell.csv")

```

IF measure clinical triple-negative one as reference
```{r IF analysis}
#This one was TNBC based on FISH for HER2 but the IHC data says HER2 2+
ref = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/Tripleneg_question/Segmentation/Quantifying_clinical_TN_p53_clin/measurements.csv")

mean(ref[Name == "Tumor",]$`Cell: Cy7 mean`)
mean(ref[Name == "Stroma",]$`Cell: Cy7 mean`)
max(ref[Name == "Tumor",]$`Cell: Cy7 mean`)

ggplot(ref, aes(x = log(`Cell: Cy7 mean`))) + geom_density()
ggplot(ref[Name == "Tumor",], aes(x = log(`Cell: Cy7 mean`))) + geom_density()
ggplot(ref[Name == "Stroma",], aes(x = log(`Cell: Cy7 mean`))) + geom_density()

ref$log_Cy7 = log(ref$`Cell: Cy7 mean`)
res <- densityMclust(ref$log_Cy7, 2)
ref$GMM_C7 = res$classification

ggplot(ref,aes(x=log_Cy7,fill = as.factor(GMM_C7)))+geom_density()

mean(ref[Name == "Tumor",]$log_Cy7)
mean(ref[Name == "Stroma",]$log_Cy7)
max(ref[Name == "Tumor",]$log_Cy7)

ref = ref[,-"ROI"]
ref = ref[,-"GMM_C7"]
ref = ref[,-"log_Cy7"]

```

```{r}
dat_all = fread("/mnt/ZTMA_21_25_26/IF_singlecell/TN_question/panCK_clin_singlecell.csv")
dat_all = dat_all[,-"Parent"]

#Merge with triple-neg ref but use only for Cy7
dat_all = rbind(dat_all,ref)

ggplot(dat_all, aes(x = log(`Cell: Cy7 mean`))) + geom_density()
ggplot(dat_all[Name == "Tumor",], aes(x = log(`Cell: Cy7 mean`))) + geom_density()

dat_all$log_Cy7 = log(dat_all$`Cell: Cy7 mean`)
res <- densityMclust(dat_all$log_Cy7, 2)
dat_all$GMM_C7 = res$classification

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/GMM.pdf")
ggplot(dat_all,aes(x=log_Cy7,fill = as.factor(GMM_C7)))+geom_density(alpha = 0.5)
dev.off()
nrow(dat_all[Image == "B038521_424_wrong_clin_p53_TN_N3.czi" & GMM_C7 == 1 & Name == "Tumor",])
dat_all$ref = 0
dat_all[Image == "B038521_424_wrong_clin_p53_TN_N3.czi", ref := 1]
pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/ref.pdf")
ggplot(dat_all,aes(x=log_Cy7,fill = as.factor(ref)))+geom_density(alpha = 0.5)
dev.off()
pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/Name.pdf")
ggplot(dat_all[Name != "Necrosis",],aes(x=log_Cy7,fill = as.factor(Name)))+geom_density(alpha = 0.5)
dev.off()

#Check spatial distribution of TN in ref
dat_all$GMM_C7 = factor(dat_all$GMM_C7,levels = c("2","1"))
png("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/spatial_TN.png")
ggplot(dat_all[Image == "B038521_424_wrong_clin_p53_TN_N3.czi" & Name == "Tumor",], aes(x = `Centroid X m`,y =`Centroid Y m`, color = GMM_C7))+geom_point(size = 0.1)+
      scale_y_continuous(trans = "reverse")+
        theme_minimal()+
      theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x= element_blank(), axis.text.y= element_blank(), axis.ticks = element_blank(),legend.position = "none")
dev.off()

#Control bar of ref
bar = dat_all[Image == "B038521_424_wrong_clin_p53_TN_N3.czi" & Name == "Tumor",]
bar[,ncells := .N,by = GMM_C7]
bar[,frac := ncells/.N]
bar = unique(bar[,c("Image","GMM_C7","frac")])
pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/ref_bar.pdf")
ggplot(bar, aes(x = Image,y = frac, fill = rev(as.factor(GMM_C7))))+geom_bar(stat = "identity")
dev.off()

#Remove ref again
dat_all = dat_all[Image != "B038521_424_wrong_clin_p53_TN_N3.czi",]

#Check background
mean(dat_all[Name == "Stroma",]$`Cell: Cy3 mean`)
mean(dat_all[Name == "Stroma",]$`Cell: Cy7 mean`)

ggplot(dat_all, aes(x = log(`Cell: Cy3 mean`))) + geom_density()
ggplot(dat_all, aes(x = log(`Cell: Cy7 mean`))) + geom_density()


#Biaxial plots
sub = as.data.table(ddply(dat,.(Image),function(x) x[sample(nrow(x),nrow(x)/10, replace = F),]))
ggplot(sub,aes(x = `Cell: Cy3 mean`,y = `Cell: Cy7 mean`))+geom_point()

#Plot density of points on scatter
ggplot(sub, aes(x = log(`Cell: Cy3 mean`),y = log(`Cell: Cy7 mean`)))+
  stat_bin2d(aes(color=..count..), bins=300, geom='point', size=1.4, fill=1)+
  scale_color_viridis(trans = "log10", option='inferno')

#Color by image
shuffle = sample(nrow(sub),replace = F)
sub = sub[shuffle,]
ggplot(sub, aes(x = log(`Cell: Cy3 mean`),y = log(`Cell: Cy7 mean`),color = as.factor(Image)))+geom_point()

#GMM on clinical markers, take in stromal cells also to have true negatives
library(mclust)

# dat_all$log_Cy7 = log(dat_all$`Cell: Cy7 mean`)
# res <- densityMclust(dat_all$log_Cy7, 2)
# dat_all$GMM_C7 = res$classification

# #Manual alternative
# dat_all[log_Cy7 <8,GMM_C7 := 1]

# ggplot(dat_all,aes(x=log_Cy7,fill = as.factor(GMM_C7)))+geom_density()
# ggplot(dat_all,aes(x=`Cell: Cy7 mean`,fill = as.factor(GMM_C7)))+geom_density()

#Also need GMM for panCK otherwise it enriches with stromal cells
dat_all$log_Cy3 = log(dat_all$`Cell: Cy3 mean`)
res <- densityMclust(dat_all$log_Cy3, 2)
dat_all$GMM_C3 = res$classification

ggplot(dat_all,aes(x=log_Cy3,fill = as.factor(GMM_C3)))+geom_density()
ggplot(dat_all,aes(x=`Cell: Cy3 mean`,fill = as.factor(GMM_C3)))+geom_density()

nrow(dat_all[Name == "Tumor" & GMM_C7 == 1 & GMM_C3 == 2,])

tumor = dat_all[Name == "Tumor",]
tumor$TN = 0
tumor[GMM_C7 == 1 & GMM_C3 == 2,TN := 1]

#Plot density of points on scatter
ggplot(tumor, aes(x = log(`Cell: Cy3 mean`),y = log(`Cell: Cy7 mean`)))+
  stat_bin2d(aes(color=..count..), bins=300, geom='point', size=1.4, fill=1)+
  scale_color_viridis(trans = "log10", option='inferno')

#Color by image
ggplot(tumor, aes(x = log(`Cell: Cy3 mean`),y = log(`Cell: Cy7 mean`),color = as.factor(GMM_C7)))+geom_point()

TN = tumor[TN == 1,]
ggplot(TN, aes(x = log(`Cell: Cy3 mean`),y = log(`Cell: Cy7 mean`),color = as.factor(Image)))+geom_point()

#Nice names for png
tumor$`Clinical marker` = "positive"
tumor[TN == 1,`Clinical marker` := "negative"]
tumor$`Clinical marker` = factor(tumor$`Clinical marker`,levels = c("positive","negative"))

#Plot distribution of negative vs positive on image
png("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/IF_TN1.png")
for (i in unique(tumor$Image)){
  print(ggplot(tumor[Image == eval(i),], aes(x = `Centroid X m`,y =`Centroid Y m`, color = `Clinical marker`))+geom_point(size = 0.1)+
      scale_y_continuous(trans = "reverse")+
        theme_minimal()+
      theme(axis.title.x=element_blank(),axis.title.y=element_blank(),axis.text.x= element_blank(), axis.text.y= element_blank(), axis.ticks = element_blank(),legend.position = "none"))
    
}
dev.off()


#Plot fraction triple-negative for each
tumor[,count := .N,by = c("TN","Image")]
tumor[,frac := count/ .N, by= "Image"]
test = unique(tumor[,c("Image","TN","frac")])
test = test[order(TN,frac),]
test[,Image := factor(Image, levels = unique(test$Image))]
g0 = ggplot(test, aes(x = Image, y = frac, fill = as.factor(TN)))+geom_bar(stat = "identity")+ theme(axis.text.x = element_text(angle = 90))

o = c("713","402","99","761","569","488","126","171","969","1364")

```

Compare to our cores
```{r}
#Get most LN met cluster fractions
met = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "lymph node mts"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t,,TissueType == "lymph node mts"))))
met = met[sizeflag != 1,]
met[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
met[,test := .N,by = "PID"]
met = met[test >=20,]
met[,frac_cluster := ncells/.N, by=c('PID')]
met = unique(met[,c('tumor_cluster_100','frac_cluster','PID')])
met = dcast.data.table(met, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)


#Get most primary tumor cluster fractions
primary = as.data.table(data.frame(colData(subset(sce_t,,TissueType == "primary tumor"))[c('PID','core','tumor_cluster_100','sizeflag')],id = colnames(subset(sce_t,,TissueType == "primary tumor"))))
primary = primary[sizeflag != 1,]
primary[,ncells:=.N ,by= c('PID', 'tumor_cluster_100')]
primary[,test := .N,by = "PID"]
primary = primary[test >=20,]
primary[,frac_cluster := ncells/.N, by=c('PID')]
primary = unique(primary[,c('tumor_cluster_100','frac_cluster','PID')])
primary = dcast.data.table(primary, 'PID ~ tumor_cluster_100',value.var = 'frac_cluster',fill = 0)

primary = primary[PID %in% o,]
met = met[PID %in% o,]

tcols = fread('/mnt/ZTMA_21_25_26/analysis/tcols.csv',header = T)
primary = melt.data.table(primary,id.vars = "PID", variable.name = "cluster",value.name = "frac")
primary[,PID := factor(PID, levels = o)]
g1 = ggplot(primary, aes(x = PID, y = frac, fill = as.factor(cluster)))+geom_bar(stat = "identity")+ theme(axis.text.x = element_text(angle = 90)) + scale_fill_manual(values = tcols$tumor_cols)

met = melt.data.table(met,id.vars = "PID", variable.name = "cluster",value.name = "frac")
met[,PID := factor(PID, levels = o)]
g2 = ggplot(met, aes(x = PID, y = frac, fill = as.factor(cluster)))+geom_bar(stat = "identity")+ theme(axis.text.x = element_text(angle = 90)) + scale_fill_manual(values = tcols[cluster %in% met$cluster,]$tumor_cols)

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/compareIMCIF.pdf")
g2 / g1 / g0
dev.off()

```

IF analysis CD11c question p53 = Cy3 and CD11c = Cy7
```{r, eval = F}
dat1 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/CD11c_question/Segmentation/p53_CD11c_first4/measurements_first4.csv")
dat2 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/CD11c_question/Segmentation/p53_CD11c_second3/measurements_second3.csv")
dat3 = fread("/home/ubuntu/tmp/server_homes/janaf/Data/2021/IF_whole_slides/CD11c_question/Segmentation/p53_CD11c_last3/measurements_last3.csv")

#Combine
dat = rbind(dat1,dat2,dat3)

#Save out IF single-cll data
fwrite(dat,"/mnt/ZTMA_21_25_26/IF_singlecell/CD11c_question/CD11c_p53_singlecell.csv")

```

Quantify number of p53 and CD11c positive cells
```{r}
dat = fread("/mnt/ZTMA_21_25_26/IF_singlecell/CD11c_question/CD11c_p53_singlecell.csv")

dat$log_Cy7 = log(dat$`Cell: Cy7 mean`)
dat$log_Cy3 = log(dat$`Cell: Cy3 mean`)
res <- densityMclust(dat$log_Cy7, 2)
dat$GMM_C7 = res$classification

ggplot(dat,aes(x=log_Cy7,fill = as.factor(GMM_C7)))+geom_density()
ggplot(dat,aes(x=log_Cy7,fill = as.factor(Name)))+geom_density()


#GMM for p53
res <- densityMclust(dat$log_Cy3, 2)
dat$GMM_C3 = res$classification

ggplot(dat,aes(x=log_Cy3,fill = as.factor(GMM_C3)))+geom_density()
ggplot(dat,aes(x=log_Cy3,fill = as.factor(Name)))+geom_density()

dat$CT = "rest"
dat[GMM_C7 == 2 & Name == "Stroma",CT := "CD11c+"]
dat[GMM_C3 == 2 & Name == "Tumor",CT := "p53+"]


#Plot distribution of cell types on image
for (i in unique(dat$Image)){
  print(ggplot(dat[Image == eval(i),], aes(x = `Centroid X m`,y =`Centroid Y m`, color = as.factor(CT)))+geom_point(size = 0.1)+
      scale_y_continuous(trans = "reverse") + ggtitle(eval(i)))
}


#Quantify numbers pos or neg
dat[,ncells := .N, by = c("CT","Name","Image")]
dat[,frac := ncells/.N,by = c("Name","Image")]

#sub = dat[(CT == "p53+" & Name == "Tumor") | (CT == "CD11c+" & Name == "Stroma"),]
sub = unique(dat[,c("Image","CT","Name","frac")])

library(stringr)
sub[str_detect(Image, "pos"),type :="CD11c+Node-"]
sub[str_detect(Image, "neg"),type :="CD11c-Node+"]
sub = sub[order(type),]
sub$Image = factor(sub$Image,levels = unique(sub$Image))

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/p53_CD11c.pdf")
ggplot(sub, aes(x = Name, y = frac, fill = CT))+geom_bar(stat = "identity")+
  facet_wrap(.~Image)
dev.off()

pdf("/home/ubuntu/tmp/server_homes/janaf/Data/2020/ZTMA21_25_26/R_dat/boxplots.pdf")
ggplot(sub, aes(x = type, y = frac, color = CT))+geom_boxplot()
dev.off()

#Correlation
c = sub[CT != "rest",]
c = dcast.data.table(c, "Image + type ~ CT", value.var = "frac")

c = rbind(c, data.table(Image = "B038935_420_redone_neg.czi",type = "CD11c-Node+",(c[Image == "B038935_420_redone_neg.czi - Scene #0",c("CD11c+","p53+")] + c[Image == "B038935_420_redone_neg.czi - Scene #1",c("CD11c+","p53+")])/2))
c = rbind(c, data.table(Image = "B987625_1433_neg.czi",type = "CD11c-Node+",(c[Image == "B987625_1433_neg.czi - Scene #0",c("CD11c+","p53+")] + c[Image == "B987625_1433_neg.czi - Scene #1",c("CD11c+","p53+")])/2))
c = c[!str_detect(Image, "Scene"),]

pos = c[type == "CD11c+Node-",]
pos = pos[Image != "B9522308_1771_redone_pos.czi",]
neg = c[type == "CD11c-Node+",]
cor(pos[,c("CD11c+","p53+")])
cor(neg[,c("CD11c+","p53+")])
```
